\bf Causal Inference in Population Trends: Searching for Demographic Anomalies in Big Data

### 2020-01-29 MANUSCRIPT/manuscript.Rmd
After Stephanie's suggested edits. Some spell checking too.

abstract: "The proliferation of big data, wider access to advanced computing platforms, and the development of powerful statistical algorithms can uncover hidden anomalies in demographic data that have previously been dismissed as noise. Here, we demonstrate how social scientists can use causal inference techniques and abductive reasoning to identify fertility and mortality anamolies in state-level data trends by investigating demographic time series data since 1999. In demonstrating our technique, we uncover real spikes and dips in fertility and mortality over time, distinguished from regular trend variations. We show that 36 states exhibited at least one mortality anomaly over time, totalling nearly 300,000 additional deaths, while an additional seven states exhibited periods of decline resuling in 250,000 fewer deaths. Ten states exhibited real, marked periods of declines in fertility, totalling more than 214,000 fewer births than expected, and 8 states had periods of excess births, resulting in an addition 12,000 births. Our results show how researchers can detect demographic anomalies in trend data and use those anomalies as a beginning point for further investigation into the causes."

>>>>

abstract: "The proliferation of big data, wider access to advanced computing platforms, and the development of powerful statistical algorithms can uncover hidden anomalies in demographic data that have previously been dismissed as noise. Here, we demonstrate how social scientists can use causal inference techniques and abductive reasoning to identify fertility and mortality anomalies in state-level data trends by investigating demographic time series data since 1999. In demonstrating our technique, we uncover real spikes and dips in fertility and mortality over time, distinguished from regular trend variations. We show that 36 states exhibited at least one mortality anomaly over time, totaling nearly 300,000 additional deaths, while an additional seven states exhibited periods of decline resulting in 250,000 fewer deaths. Ten states exhibited real, marked periods of declines in fertility, totaling more than 214,000 fewer births than expected, and 8 states had periods of excess births, resulting in an addition 12,000 births. Our results show how researchers can detect demographic anomalies in trend data and use those anomalies as a beginning point for further investigation into the causes."

### 2020-01-29 MANUSCRIPT/manuscript.tex
After Stephanie's suggested edits. Some spell checking too.

\bigskip
\begin{abstract}
The proliferation of big data, wider access to advanced computing
platforms, and the development of powerful statistical algorithms can
uncover hidden anomalies in demographic data that have previously been
dismissed as noise. Here, we demonstrate how social scientists can use
causal inference techniques and abductive reasoning to identify
fertility and mortality anamolies in state-level data trends by
investigating demographic time series data since 1999. In demonstrating
our technique, we uncover real spikes and dips in fertility and
mortality over time, distinguished from regular trend variations. We
show that 36 states exhibited at least one mortality anomaly over time,
totalling nearly 300,000 additional deaths, while an additional seven
states exhibited periods of decline resuling in 250,000 fewer deaths.
Ten states exhibited real, marked periods of declines in fertility,
totalling more than 214,000 fewer births than expected, and 8 states had
periods of excess births, resulting in an addition 12,000 births. Our
results show how researchers can detect demographic anomalies in trend
data and use those anomalies as a beginning point for further
investigation into the causes.
\end{abstract}

>>>>

\bigskip
\begin{abstract}
The proliferation of big data, wider access to advanced computing
platforms, and the development of powerful statistical algorithms can
uncover hidden anomalies in demographic data that have previously been
dismissed as noise. Here, we demonstrate how social scientists can use
causal inference techniques and abductive reasoning to identify
fertility and mortality anomalies in state-level data trends by
investigating demographic time series data since 1999. In demonstrating
our technique, we uncover real spikes and dips in fertility and
mortality over time, distinguished from regular trend variations. We
show that 36 states exhibited at least one mortality anomaly over time,
totaling nearly 300,000 additional deaths, while an additional seven
states exhibited periods of decline resulting in 250,000 fewer deaths.
Ten states exhibited real, marked periods of declines in fertility,
totaling more than 214,000 fewer births than expected, and 8 states had
periods of excess births, resulting in an addition 12,000 births. Our
results show how researchers can detect demographic anomalies in trend
data and use those anomalies as a beginning point for further
investigation into the causes.
\end{abstract}
